### 2025-01-14 technical_doc/design.tex
fix labels, fix problem with graphics path and suround with curly braces, start technical document

% !TEX root = ./skripsi.tex
\chapter{RESULTS AND DISCUSSION}
\noindent This chapter will elaborate on the results of the study and relevant phases for analysis. This chapter is organized into five sections which are the results of data generation, data retrieval, design of computational model, implementation of computational model, experimental scenarios, and the experiments themselves.

\section{Data Generation}\label{sec:data_generation}
\noindent This study uses two data acquisition approaches. The first approach which is discussed in this section is data generation. This process involves manufacturing data randomly in such a way that they obey the PDEs to be modeled. The generated data consists of features and labels. These features and labels in the context of PDEs are parameters and the solution of the PDEs, respectively. The PDEs enforce a relationship between parameters and the solution. This relationship is implicitly encoded into the features and labels, which is what machine learning models can learn. Because the solution and parameters to PDEs are functions and there are many kinds with different properties, generating all the different kinds of functions is a difficult task. This is why this study focuses on Fourier functions. This means that out of the space of all functions \(F \) which include categories such as polynomials \(f\left(x\right)=a_{n}x^n+\cdots+a_1x+a_0\) where \(f\in F\), and \(a_n\) are constants; we only consider the subset of functions \(U \subset F\) which are of the form \(u\left(x\right)=\sum_{k}^{m}c_k e^{2\pi i k \frac{x}{P}}\).

Using the subset of Fourier functions has several benefits which has motivated the choice. First these functions are characterized purely by their coefficients \(c_k\), meaning there is no need to store discretized values which potentially saves space and computation. The second benefit is that other functions such as polynomials can be approximated by them using the Fourier transform. This means that despite limiting the set of functions to Fourier functions, the behavior of PDEs with other sets of functions can be approximated to a certain extent. The final benefit is the mature ecosystem around these functions which include fast algorithms for the Fourier transform and even numerical approaches for solving known PDEs like the previously mentioned spectral method which we discussed in \lccref{sec:fourier_discussion}. In summary, the generated data consists of features and labels which are Fourier coefficients implicitly defining the relationship enforced by a PDE and scenario. To proceed with the data generation, there are 4 steps involved:
\begin{enumerate}
  \item Scenario and PDE Determination:
        The first step of data generation is determining the scenario related to the PDE or governing equations. First, the PDE to be modeled is determined based on the goals of the dataset. Then, one or more of the parameters is chosen to predict the solution. The chosen parameters will be called the input functions and the solution will be called output functions from here on. The second part of the scenario is the domain or more simply the physical space occupied by the system to be modeled. The domain will be used to compute the function values in relation to the physical space from coefficients of Fourier functions.
  \item Parameter Determination:
        The second step is determining all parameters other than that input parameter based on the scenario and governing equation. These parameters may be coefficients such as material properties like density or viscosity, or forcing terms which model external influence on the system like a heat source in the case of the heat equation. Depending on the parameters, solutions of PDEs may behave very differently, such as the appearance of discontinuities in the solution to low viscosity Burgers' equation. Because of this, the choice of parameters is guided by what the dataset seeks to do.
  \item Random Coefficient Generation:
        The third step of data generation is randomly generating the solutions for the chosen equation. Generating random functions in the space of Fourier functions exploits the fact that the coefficients characterize the function completely. By randomly assigning coefficients \(c_k\), many functions can be generated randomly with very little cost. Since the coefficients need to be complex numbers as in \lccref{eq:complex_number}, both real and imaginary components are generated independently by assigning a random value to each component for each wave number \(k\). They are then put together again into complex numbers.

        Since only real functions are of interest in this study, the generation cost can be approximately halved. This is because for real functions the coefficients for negative wave numbers \(k\) are complex conjugate of the positive wave numbers. This means that once the positive coefficients are generated, one only needs to compute their complex conjugate and concatenate the result with the coefficients of positive wave numbers. For dimensions higher than one, a simpler approach is used. The coefficients are generated for all wave numbers including the negative ones. The inverse Fourier transform is computed and this results in complex functions. The real components of these functions are kept and the Fourier transform is applied to get the coefficients of the real functions. Finally, these generated coefficients then be used with the basis functions as input functions.
        % TODO: add diagram for generating real functions, the mirroring process too
  \item Forcing Term Computation:
        Finally, in order to ensure that the generated solution and chosen parameters satisfy the equation, the forcing term is computed as the residual of the equation of interest with the parameters that was previously determined. This computation is done using the spectral method.
  \item Function Value and Noise Computation:
        The generated solution and forcing functions are labeled as input and output functions. The values of both functions in the domain can then be computed using the scenario determined in the first step. Once all coefficients are generated, the function values are computed with an inverse Fourier transform. The real component of the function values are retained, and the imaginary component are zeroed out. Noise is added to the function values here as needed. The processed function values are then converted back into coefficients using a Fourier transform. This processing is necessary to ensure that the coefficients are only describing the real function.

These four steps are the general processes involved in generating datasets for this study. Further specifics of the generation process of each dataset is explained in their respective subsections.

\subsection{Data Generation: Anti-derivative}
\noindent The first dataset is a simple one dimensional derivative. This was chosen as a simple proof of concept of the ability to solve a differential equation. The equation is related to many real-world problems such as acceleration and speed. One can imagine a train in an ideal world where acceleration is directly translated into speed. When the train accelerates at time \(t \) by some amount \(a \), we can expect the train to have some speed \(u \). In this ideal world, the relationship between speed and acceleration can be modeled with a simple differential \lccref{eq:derivative}. This scenario is found in many real systems albeit often with many more details such as different components of acceleration from friction, drag, gravity, and other factors.
As previously mentioned in \lccref{sec:data_generation}, both velocity \(u \) and acceleration \(a \) are modeled with Fourier series in \lccrefs{eq:fourier_speed,eq:fourier_acceleration} respectively, where \(T\) is the period representing the length of the domain.
\begin{align}
  u\left( t \right) & = \sum_{k} \hat{u}_k e^{2\pi ikt/T} \label{eq:fourier_speed}        \\
  a\left( t \right) & = \sum_{k} \hat{a}_k e^{2\pi ikt/T} \label{eq:fourier_acceleration}
\end{align}
The domain of the scenario is a two-hour time window. This number was chosen because it is around the ideal length of travel time on high speed rail in comparison to air travel and car travel \autocite{givoniDevelopmentImpactModern2006,wangEfficiencySpatialEquity2019,wrro2236}. This is the first step in generating this dataset.

In the second step, as the derivative \lccref{eq:derivative} does not contain any parameters other than the acceleration which is the input parameter, there are no other parameters to determine. Therefore, the data generation process proceeds to generating coefficients for the speed functions \(\hat{u} \). The coefficients are assigned randomly from a Gaussian distribution with a mean of zero and standard deviation of one. This choice was made such that most wave numbers will have a coefficient of close to zero leaving a sparse set of wave numbers to mostly affect the resulting function. In total, \num{5000} unique functions are generated with \num{50} complex coefficients each.

In the third step, the output function coefficients are computed. To find the relation between \(\hat{u}_k \) and \(\hat{a}_k \), we need to find substitutes for each term in \lccref{eq:derivative}. To do this, we take the derivative of \lccref{eq:fourier_speed} which result in \lccref{eq:fourier_series_derivative}. Using this we can substitute the terms in \lccref{eq:derivative} with \lccrefs{eq:fourier_series_derivative,eq:fourier_acceleration} giving \lccref{eq:example_spectral_method_fourier_substituted}. Finally, after some algebraic manipulation we obtain the relationship between the input \(a^*\) and output function \(u^*\) in terms of their coefficients in \lccref{eq:derivative_coeff}. One also needs to choose the integration constant \(\hat{u}_{0}\) because at \(k=0\) \lccref{eq:derivative_coeff} becomes a division by zero.
\begin{align}
  \dv{u\left( t \right)}{t}                                        & = a\left(t\right) \label{eq:fourier_series_derivative}                                                             \\
  \sum_{k} \hat{u}_k\times \left( 2\pi ik/T \right) e^{2\pi ikt/T} & = \sum_{k} \hat{a}_k e^{2\pi ikt/T} \label{eq:example_spectral_method_fourier_substituted}                         \\
  \hat{a}_k                                                        & = \hat{u}_k\left( 2\pi ik/T \right)                                                    \label{eq:derivative_coeff}
\end{align}

In the final step, using randomly generated values of \(\hat{a}_k\), the corresponding values of \(\hat{u}_k\) are computed with \lccref{eq:derivative_coeff}. These coefficients are then used to compute the function values inside the domain. The two-hour time window is represented by a grid of \num{500} discrete points. With the coefficients and evaluation points ready, the function values are computed using \lccrefs{eq:fourier_speed,eq:fourier_acceleration}. Next, noise is added to the function values in order to motivate models learning on the dataset to generalize on noise. To also allow evaluation of how well the model performs with different levels of noise, the samples are duplicated into three copies for each high, medium and low noise levels. The function values of each copy is perturbed with noise from a Gaussian distribution with zero mean and standard deviation of some percentage of the average function value standard deviation. The percentages of each high, medium, and low noise levels are 5, 10, and 50 percent. The perturbed function values are then transformed back into their coefficients. An example generated function and its different perturbed versions is shown in \lccref{fig:antiderivative_noise_levels}. The horizontal axis indicates time which is displayed in units of hours. The vertical axis indicates velocity in abstract units.

\begin{figure}[H]
  \centering
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_clean_perturbed_solution.pgf}
    \end{adjustbox}
    \caption{}\label{fig:antiderivative_noise_levels}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_clean.pgf}
    \end{adjustbox}
    \caption{}\label{fig:antiderivative}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/antiderivative_derivative_clean.pgf}
    \end{adjustbox}
    \caption{}\label{fig:derivative}
  \end{subfigure}
  \caption{(\subref{fig:antiderivative_noise_levels}) A generated function with no noise (clean), low noise level (5\%), medium noise level (10\%), and high noise levels (50\%).\ (\subref{fig:antiderivative}) Antiderivative function with no noise.\ (\subref{fig:derivative}) Derivative function with no noise.}
\end{figure}
% TODO: interesting note that the model trains better on function to coefficient task when the antiderivative is computed from the randomly generated derivative than if the derivative is computed from the randomly generated antiderivative. THIS NEEDS A REASON AND ANALYSIS SO MIGHT NOT BE NEEDED IN THIS WRITING. MAYBE A GOOD IDEA FOR A SEPARATE PIECE

\subsection{Data Generation: Burgers' Equation}\label{sec:data_generation_burgers}
\noindent The second dataset generated in this study concerns the Burgers' equation. This equation has been used to model a variety of cases including fluid dynamics, traffic flow, and, shock waves \autocite{bonkileSystematicLiteratureReview2018,orlandiBurgersEquation2000,becBurgersTurbulence2007,jamesonEnergyEstimatesNonlinear2007}. This equation is also used as a base problem for testing the effectivity of numerical methods in solving non-linear PDEs \autocite{barterShockCapturingPDEbased2008,banksNumericalErrorEstimation2012, tabatabaeiImplicitMethodsNumerical2007,bonkileSystematicLiteratureReview2018}. The nonlinear term in the equation creates steep gradients and even shock waves which are discontinuous with low viscosity conditions. These challenges test the stability of numerical solvers. This is the scenario and reason for the choice of this PDE\@. To control the viscosity and therefore the steepness of gradients, the formulation of the Burgers' equation considered in this study is the forced viscous Burgers' equation in one dimension. For a velocity of \(u(x, t)\), viscosity of \(\nu \), and forcing term of \(f(x, t)\), the formulation of the forced Burgers' equation is shown in \lccref{eq:forced_viscous_burgers}.

The domain of we consider here is based on an exact solution of the equation. This is done so that the learned model can be tested on an exact solution and capture enough of the details within the solution within the domain. In this study, we use a specific solution by \textcite{woodExactSolutionBurgers2006} which can be seen in \lccref{eq:burgers_exact_solution}. This solution is periodic in space but not in time. Therefore, the domain in space will be based on the spatial periodicity of the exact solution which is two. As such, the space domain spans from 0 to 2. The time domain, on the other hand, is chosen to span from 0 to 10 such that enough details of the solution is included.

Continuing to the second step, the parameters involved in \lccref{eq:forced_viscous_burgers} are the viscosity \(\nu \) and the forcing term \(f \). Since the forcing term is dependent on the solution in this step, the only parameter to determine is the viscosity. We choose three constant viscosity values which are \num{0.1}, \num{0.01}, and \num{0.0}. These values are chosen so that a variety of behaviors are represented in the dataset from viscous to inviscid flow.

After this, in the third step, the solution functions \(u \) are generated with 8 modes in time and 8 in space and \num{500} unique function samples for each viscosity value. Since we don't factor the viscosity in at this point, for ease of comparison, we ensure the same solutions are generated every time by using a new instance of the generator with the same seed number each time. This results in the solutions being the same for all viscosity and other parameters of the equation are adapted to this.

Next we account for the different viscosity values in this stage. The forcing term is computed using \lccref{eq:forced_viscous_burgers}. The solution field \(u\) and forcing term \(f\) are modeled with Fourier series in \lccrefs{eq:fourier_field,eq:fourier_force} where \(k\) is a vector of spatial and temporal wave number such that \(k_t\) is the temporal wave number and \(k_x\) is the spatial wave number. The period in time is represented by \(T\) and in space it is \(L\).
\begin{align}
  u\left(x, t \right) & = \sum_{k_{x}} \sum_{k_{t}} \hat{u}_k e^{2\pi i(k_{x}x/L+k_{t}t/T)} \label{eq:fourier_field} \\
  f\left(x, t \right) & = \sum_{k_{x}}\sum_{k_{t}} \hat{f}_k e^{2\pi i(k_{x}x/L+k_{t}t/T)} \label{eq:fourier_force}
\end{align}
Substituting the terms in \lccref{eq:forced_viscous_burgers} with the respective Fourier series, we get the equation \lccref{eq:forced_viscous_burgers_fourier_substitution}. A problem one notices is the nonlinear term \(u\pdv{u}{x}\). A naive approach would multiply all wave numbers terms with each other as shown in \lccref{eq:burgers_convolution}.
\begin{align}
  u\pdv{u}{x} = \left(\sum_{k_{x}} \sum_{k_{t}} \hat{u}_k e^{2\pi i(k_{x}x/L+k_{t}t/T)}\right) \times \left(\sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{x}/L)\hat{u}_k e^{2\pi i(k_{x}x/L+k_{t}t/T)}\right)\label{eq:burgers_convolution}
\end{align}
This is computationally expensive operation with \(N^2\) multiplications \autocite{lariosMATH934BURGERS2021,robertsDealiasedConvolutionsPseudospectral2011,shenSpectralMethodsAlgorithms2011,orszagComparisonPseudospectralSpectral1972}. To avoid this complexity, the term is first reformulated into \(\pdv{x}(u^2/2)\). Then to avoid aliasing, the coefficients are padded with 50\% zeros such that the padded coefficients are 3/2 times the size of the original \autocite{orszagEliminationAliasingFiniteDifference1971, lariosMATH934BURGERS2021}. Then we use the padded coefficients to compute the function values in the physical domain using the inverse transform. And then, the point wise squaring operation is performed and transform the results back to spectral domain. The resulting coefficients are then trimmed back to their original size before padding. Finally, we multiply the resulting coefficients \(\hat{uu}_k\) by the derivative constants. This operation is much more efficient with only \(1.5 N\) multiplications. This is still more efficient than the naive approach even after taking into account the transforms involved adds \(O(2\times 1.5Nln(1.5N))\) operations if using a Fast Fourier Transform algorithm. Putting all the above together, results in \lccref{eq:forced_viscous_burgers_fourier_substitution}.
\begin{equation}
  \begin{split}
    \sum_{k_{x}} \sum_{k_{t}} \hat{f}_k e^{2\pi i(k_{x}x/L+k_{t}t/T)} =
     & \sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{t}/T) \hat{u}_k e^{2\pi i(k_{x}x/L+k_{t}t/T)}                                                                \\
     & + \sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{x}/L)\hat{uu}_k e^{2\pi i(k_{x}x/L+k_{t}t/T)}                                                              \\
     & - \nu\sum_{k_{x}} \sum_{k_{t}} {(2\pi i k_{x}/L)}^2\hat{u}_k e^{2\pi i(k_{x}x/L+k_{t}t/T)} \label{eq:forced_viscous_burgers_fourier_substitution}
  \end{split}
\end{equation}
\begin{align}
  \sum_{k_{x}} \sum_{k_{t}} \hat{f}_k & = \sum_{k_{x}} \sum_{k_{t}} (2\pi i k_{t}/T) \hat{u}_k + (2\pi i k_{x}/L)\hat{uu}_k - \nu{(2\pi i k_{x}/L)}^2\hat{u}_k \label{eq:forced_viscous_burgers_fourier} \\
  \hat{f}_k                           & = (2\pi i k_{t}/T) \hat{u}_k + (2\pi i k_{x}/L)\hat{uu}_k - \nu{(2\pi i k_{x}/L)}^2\hat{u}_k \label{eq:forced_viscous_burgers_coeff}
\end{align}
After simplifying the equation, we get the coefficients as in \lccref{eq:forced_viscous_burgers_coeff}. Using this equation, the exact forcing term corresponding to the randomly generated solution can be computed with relatively low cost. Finally, the function values of both the solutions and forcing terms are perturbed by adding Gaussian noise with a mean of zero and standard deviation of 10\% of the function value standard deviation to the values of the inverse transform. The perturbed coefficients are then recomputed from the sum and the final perturbed functions are obtained.

An example of the solution and forcing term pairs that has been generated is shown in \lccref{fig:burgers_data}. Using the same seed we can see that the solution functions of the same sample are the same. However, with the forcing terms, we see much larger amplitudes for higher frequencies for larger viscosity values. This is inline with \lccref{eq:forced_viscous_burgers_coeff}. The squared multiplier gets large much more quickly for higher frequencies due to the squaring operation. The since this effect is controlled by the viscosity via multiplication, the higher the viscosity, the larger the high frequency amplitudes of the forcing term. This, in other words, means that more viscous fluids require a stronger forcing term to affect the solution in the same way with as fluids with a lower viscosity.

\begin{figure}[H]
  \centering
  \begin{subfigure}{0.49\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/burgers_solution_0.0.pgf}
    \end{adjustbox}
    \caption{Solution function for \(\nu=0\).}\label{fig:burgers_solution_0.0}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/burgers_forcing_0.0.pgf}
    \end{adjustbox}
    \caption{Forcing function for \(\nu=0\).}\label{fig:burgers_forcing_0.0}
  \end{subfigure}
  % \\[-0.7\baselineskip]
  \begin{subfigure}{0.49\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/burgers_solution_0.01.pgf}
    \end{adjustbox}
    \caption{Solution function for \(\nu=0.01\).}\label{fig:burgers_solution_0.01}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/burgers_forcing_0.01.pgf}
    \end{adjustbox}
    \caption{Forcing function for \(\nu=0.01\).}\label{fig:burgers_forcing_0.01}
  \end{subfigure}
  % \\[-0.7\baselineskip]
  \begin{subfigure}{0.49\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/burgers_solution_0.1.pgf}
    \end{adjustbox}
    \caption{Solution function for \(\nu=0.1\).}\label{fig:burgers_solution_0.1}
  \end{subfigure}
  \begin{subfigure}{0.49\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/burgers_forcing_0.1.pgf}
    \end{adjustbox}
    \caption{Forcing function for \(\nu=0.1\).}\label{fig:burgers_forcing_0.1}
  \end{subfigure}
  \caption{Example sample pairs of solution and forcing terms for the Burgers' equation dataset.}\label{fig:burgers_data}
\end{figure}

\section{Data Retrieval}\label{sec:data_retrieval}
\noindent The second data acquisition approach retrieves data of a from an external source. This approach downloads the external data to a local machine. In this study, the specific dataset that will be used is the ERA5 dataset from the European Center for Medium-range Weather Forecast (ECMWF). Specifically, the ERA5 dataset is a reanalysis which means it combines observational data from all over the world in order to present a more complete picture of the weather system. This combination process, named 4D variational data assimilation, takes into account the physics known to be involved in the system. As an example weather station data which take measurements such as wind speed, humidity, and temperature of the immediate surroundings of the weather station is very limited in giving a broader picture of weather over a larger area. Even technologies like earth observations satellites are limited to what they can observe at any single point in time as most cannot see the entirety of the earth's surface at once. The ERA5 dataset specifically, assimilates previous forecasts with observational data every 12 hours. The ERA5 reanalysis is available at the following URL (\url{https://cds.climate.copernicus.eu/datasets/reanalysis-era5-single-levels}).

The original dataset made available by ECMWF has a grid size of \ang{0.25} by \ang{0.25} in latitude and longitude for atmospheric data. Data of oceanic waves are also available at a coarser grid size of \ang{0.5} by \ang{0.5}. However, we will use a cloud optimized version called WeatherBench2 that is easier to retrieve \autocite{raspWeatherBench2Benchmark2023}. This is because the data is available in several more grid sizes such as \ang{1.5}. In addition, the data can be readily downloaded to a local machine without waiting for further processing on the server. The guide for working with the dataset is available at (\url{https://weatherbench2.readthedocs.io/en/latest/data-guide.html}). In our specific case, we use the version labeled \verb|1959-2023_01_10-6h-64x32_equiangular_conservative.zarr|. This dataset spans from year 1959 to 2023 with a grid size of \ang{5.625} or resolution of 64 by 32 which spans from \ang{-87.19} to \ang{87.19} in latitude and \ang{0.0} to \ang{354.4} in longitude. The dataset is also resampled down to 6-hour intervals that starts at midnight UTC January 1959 1\textsuperscript{st} and ends at 18:00 UTC January 10\textsuperscript{th} 2023.

While the original data provided by ECMWF is available in NetCDF or GRIB, the WeatherBench2 version is provided using the Zarr format in a Google Cloud Storage Bucket. This setup allows the use of libraries that support the Zarr format to access the dataset remotely. The particular library we use is called \emph{Xarray} version \verb|2024.9.0| \autocite{hoyer2017xarray,hoyerXarray2024}. Accessing the remote dataset is done as shown in \lccref{fig:access_remote_zarr}. In order to ensure access is granted, the storage option token is set to anon so that the method uses the anonymous only public access mode of authentication \autocite{GCSFSGCSFs2023122post1+1g8e500c6dirty}.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
  import xarray as xr
  ds = xr.open_zarr(
      "DATASET_ADDRESS",
      storage_options = {"token": "anon"},
  )
  \end{lstlisting}
  \caption{Example of using Xarray to access a publicly accessible Zarr dataset hosted on Google Cloud Storage Bucket.}\label{fig:access_remote_zarr}
\end{figure}

The specific dataset resolution we use is 64 by 32 representing longitude and latitude respectively. The ERA5 reanalysis dataset is hosted at the following URL\@: \url{gs://weatherbench2/datasets/era5/1959-2023_01_10-6h-64x32_equiangular_conservative.zarr}. The corresponding climatology is hosted at \url{gs://weatherbench2/datasets/era5-hourly-climatology/1990-2019_6h_64x32_equiangular_conservative.zarr}. The climatology is the average for hours 0, 6, 12, and 18 for each day of the year from 1990 until 2019. This average helps in defining the expected state of the weather at specific points in time based on the historical averages for that time of day and year.

For the reanalysis dataset, we limit the time range to the years 1995{-}2013 for training and 2020{-}2022 for testing. The specific time ranges for training are chosen to balance the diversity within the data and capacity of the machine the processing will be done on. The testing set time range was chosen to be as close to those shown on the WeatherBench 2 website. The filter for time and variables is shown in \lccref{fig:wb2_retrieval}. The reason for the different time ranges between training and testing is to reduce the risk of information leakage across the training and testing sets. This ensures that the testing scores we measure will be for the most part from the generalization that model learned during training. The filter is an array the length of the number of samples. The array values are all the datetime values that the year ranges are true for. This results in an array of boolean values. Using this boolean filter we can index the data so that we get the filtered data. Then, the filtered data is saved locally so that we can access it in a more performant way. The encoding chunks may need to be removed in order to evade the Xarray library from mistakenly persisting the remote encoding after slicing data \autocite{ianZarrEncodingAttributes}.
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
  years_idx = ds['time.year']
  data_idx = (years_idx > START_YEAR) & (years_idx < END_YEAR)
  da_orig = ds['2m_temperature'][data_idx]
  # save locally
  da_orig.encoding.pop('chunks',None)
  da_orig.to_zarr(DATA_FILENAME, mode='w')
  # open local version
  da = xr.open_zarr(DATA_FILENAME)
  \end{lstlisting}
  \caption{Example of filtering of Xarray dataset by year and variable. The data is saved and then loaded locally.}\label{fig:wb2_retrieval}
\end{figure}

The dataset includes 62 variables with some variables also spanning 13 discrete vertical levels in the atmosphere. For our use, we will only be using the 2-meter temperature of the atmosphere. This is the air temperature 2 meters above the surface be it land, sea, or inland water. This variable is in units of kelvin (K). Vertically, the values of this variable is obtained by interpolating between the model values at the surface and the lowest model vertical level. An example of the data that was retrieved can be seen in \lccref{fig:wb2_data}. The data shown is in the equiangular projection. This means that towards the poles, the shapes become more distorted. While this is not the most ideal way to represent the data, using the data in this way adds to the challenge of how the model can adapt to the use of suboptimal configurations in the data. To compute the anomaly, we can simply subtract the current state in \lccref{fig:wb2_train_data} by the climatology in \lccref{fig:wb2_clim_data} resulting in \lccref{fig:wb2_train_anomaly}.
\begin{figure}[H]
  \centering
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/wb2_train_data.pgf}
    \end{adjustbox}
    \caption{2-meter temperature}\label{fig:wb2_train_data}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/wb2_clim_data.pgf}
    \end{adjustbox}
    \caption{Climatology}\label{fig:wb2_clim_data}
  \end{subfigure}
  \begin{subfigure}{\linewidth}
    \begin{adjustbox}{width=\linewidth}
      \input{figures/wb2_train_anomaly.pgf}
    \end{adjustbox}
    \caption{2-meter temperature anomaly}\label{fig:wb2_train_anomaly}
  \end{subfigure}
  \caption{Data retrieved: (\subref{fig:wb2_train_data}) 2-meter temperature at January 1, 1995 6:00 UTC\@. (\subref{fig:wb2_clim_data}) 2-meter temperature climatology for the first day of the year at hour 6 UTC\@. (\subref{fig:wb2_train_anomaly}) Anomaly for 2-meter temperature at January 1, 1995 6:00 UTC\@.}\label{fig:wb2_data}
\end{figure}

After filtering, our training set has the dimensions of \num{27760} time steps, \num{64} longitude, and \num{32} latitude. The testing set on the other hand has \num{4384} time steps and the same values for the other two dimensions. For uses, such as a validation set during hyperparameter optimization, one can use a subset of the training set. The dimensions of the climatology are 4 for hour (0, 6, 12,\& 18), \num{366} for day of year, \num{64} for longitude, and 32 for latitude.

\section{Design of Computational Model}\label{sec:computational_model_design}
\noindent In this section, we will discuss the design of the computational model that is proposed in this work. An illustration of the computational model is show in \lccref{fig:computational_model}.

\begin{figure}[H]
  \centering
  \begin{adjustbox}{max width=\textwidth}
    \tikzfig{figures/computational_model}
  \end{adjustbox}
  \caption{Computational Model of SpectralSVR.}\label{fig:computational_model}
\end{figure}

The computational model is divided into four large phases. The processes within these phases are based on the broad outcomes of the phase. The phases themselves are data preparation, data transformation, model training, and model evaluation. Each phase is composed of processes that together works toward the output of the larger process. At the end, a final process of analysis and interpretation is carried out.

\subsection{Data Preparation}
\noindent The data preparation phase starts off the computation by preparing the data that is to be used in this study. This phase starts with specifying the problem to be solved. This translates to the dataset that will be used to represent these problems. The first type of dataset is generated based on a known governing equation or PDE\@. Using said equation, the goal is to generate pairs of functions that are related to each other by the equation. In the next process, generally, the solution function is first generated randomly. This is done by generating random numbers to use as coefficients of a set of basis functions. Then parameters of the equation is determined, such as term coefficients. Using the generated solution and predetermined parameters, a forcing term can generally be computed to satisfy the equation. Augmentations such as noise can then be added in the final process which is evaluating the function values. These processes result in a generated dataset. To verify the generation method and any models trained on the generated data one can use exact analytical solutions of the governing equations.

The second type of dataset is retrieved from an external source. The data source is first identified, in our case it is the WeatherBench 2 collection of datasets. The specific dataset from the collection of different dataset versions is first determined. The dataset is available in the \emph{Zarr} format. And then, we determine the filters for the portions of the dataset we are interested in. Once the dataset and filters are determined, the retrieval process starts by accessing the dataset using the \emph{Xarray} library \autocite{hoyer2017xarray,hoyerXarray2024}. The filters are applied as in \lccref{fig:wb2_retrieval}. Finally, the filtered version is stored locally. The local dataset is also stored in the \emph{Zarr} format.

\subsection{Data Transformation}\label{sec:data_transformation}
\noindent The second phase of the computational model transforms the data that has been prepared into a form that can be used with the proposed model. In this study, there are two ways a dataset can be used. The first and simplest way is to predict the solution function from the forcing term for a given set of PDE parameters. In this case, the input function is the forcing term. In other words, the input function is the predictor or independent variable that will be used to predict the solution function. As such, the output function is the solution function. The antiderivative dataset is used in this way. The derivative, which represents the acceleration function, is used to predict the velocity function. Which means that the input function is the acceleration function and the output function is the velocity function.

The second use of a dataset depends on whether it can be interpreted as a mapping of a previous function to the next function. In other words, a series of functions related to the previous function as a sum of the previous function and a perturbation. A commonly used form of this are time series. Generally, a time series can be represented as \lccref{eq:time_step_function}. This function describes the evolution of function \(f\) as a sum of some initial state \(f^0\) and perturbation which is its rate of change \(\dv{f}{t}\) at a point in time close to the current time multiplied by the time step \(\delta t\). This is the first order Taylor polynomial. Adding more terms with higher order derivatives can be done to increase the accuracy. This describes the same process that is used in the finite difference method.
\begin{equation}
  f^{t+1} = f^{t} + \delta t \times \dv{f}{t} + \dots \label{eq:time_step_function}
\end{equation}
In this case the previous function and any other independent variables like the forcing term involved in the right-hand side of the equation is used to predict the difference between the future state and the current state. It would mean that each sample function in the data would need to at least have two dimensions. This is because one dimension represents the evolution of the function over time, while the other dimensions define the spatial domain where the function's values are computed at each time step. This condition is true for both the Burgers' equation dataset and the WeatherBench 2 collection of ERA5 datasets. First, the Burgers' equation dataset, while is also able to be used in the first way as a mapping between the forcing term and the solution, one of the traditional ways of solving it is using the method of lines. This entails solving the dimensions other that time with methods such as the finite differences method or the spectral method. The time dimension is then separately solved using FDM such as the form shown in \lccref{eq:time_step_function} \autocite{schiesserNumericalMethodLines2012,sadikuSimpleIntroductionMethod2000}. This is the form that the Burgers' equation dataset will be used in. In this case, algebraic manipulation of \lccref{eq:forced_viscous_burgers} results in the time derivative of the solution as in \lccref{eq:forced_viscous_burgers_time_derivative}.
\begin{align}
  \pdv{u}{t} = f - u\pdv{u}{x} + \nu\pdv[2]{u}{x} \label{eq:forced_viscous_burgers_time_derivative}
\end{align}
The most important component of the right-hand side of the equation, the time derivative, is expressed in terms of the solution function, the forcing term and any other partial derivatives. Because of this, it means we can train the model to learn the relationship between the solution function combined with the forcing term and the time derivative. Putting this together with \lccref{eq:time_step_function}, it is possible to predict the solution function in a future time based on the current solution function and forcing term. The input functions in this case are both the value of the solution function and the forcing term at the current time step. The output function is the time derivative which is computed by rearranging \lccref{eq:time_step_function} resulting in \lccref{eq:time_derivative_time_step_function}.
\begin{equation}
  \dv{f}{t} = \frac{f^{t+1} - f^{t}}{\delta t} \label{eq:time_derivative_time_step_function}
\end{equation}
Perhaps in a more obvious manner, the same is also applied to the WeatherBench 2 collection of ERA5 datasets. The 2-meter temperature is a series of snapshots in time of the distribution of temperature all over the world at 2 meters above the surface. The input function in this case is the temperature distribution at the current time step. The output function is the time derivative of the temperature distribution computed using \lccref{eq:time_derivative_time_step_function}.

The input and output functions are then transformed into basis coefficients using the Fourier transform we discussed in \lccref{sec:fourier_discussion}. Once the coefficients are computed, the input and output sample pairs are split into train, test, and validation subsets. For the first type of input and output function pair, since the samples are independent of each other, a random sampling of the dataset without replacement is done to select the samples for each subset. For the second type of input and output function pairs, if there are multiple samples of functions as in the Burgers' equation dataset, the same process with random sampling is followed. On the other hand, if only one function sample is available, such as the WeatherBench 2 dataset, an issue that arises is the fact that the samples are dependent on one another. This is because a previous time step influences the next time step. This is a problem for generalizing the learned model, since random sampling would lead to the model learning information in the future of the testing subset \autocite{kapoorLeakageReproducibilityCrisis2023,kaufmanLeakageDataMining2012}. As such, the portioning of the dataset is done by dividing the sample into different sections of the function in time.

>>>>

% !TEX root = ./technical_doc.tex
\chapter{DESIGN}\label{sec:implementation_design}
\noindent In this subsection, the architectural design of the software product is explained. The architecture consists of modules with similar concerns. This was chosen because of the intuition built by the use of other libraries with a similar architecture. The broad aim of modeling PDEs with LSSVR and basis functions coefficients can be separated into smaller concerns of data, model, and basis functions. Each of these concerns are addressed by a corresponding module. The explanation of each module are as follows:
\begin{enumerate}
  \item Utilities
